---
title: "Predicting Player Rating in FIFA"
author: "Junqi Liao, Kaiwen Dong, Raymond Tan"
date: "07/08/2019"
output: pdf_document
toc: yes
toc_section: 2
theme: united
highlighted: tango
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/Raymond/Desktop/Raymond Tan/HW/4B/STAT444/soccer-rating-prediction/data")
soccer.raw <- read.table("rating_potential.csv",sep = " ",na.strings = "NA")
library(glmnet)
library(data.table)
library(ggplot2)
library(MASS)
library(rpart)
library(rpart.plot)
library(randomForest)
set.seed(123)
```




# Motivation and introduction of Report

# Data

# Preprocessing

## Data Visualization


# Smoothing Methods
## Linear Models
### Multiple Linear Regression

### LASSO Regression

### Ridge Regression

```{r,echo=FALSE}
mse <- function(y,yhat) {
  return( mean( (y - yhat)^2 ))
}


```

```{r,echo=FALSE}
soccer <- soccer.raw
soccer$player_name <- NULL
soccer$set <- ifelse(runif(n=nrow(soccer)) > 0.85,yes = 2,no = 1)
#Split data into training set and testing set
soccer.train <- soccer[which(soccer$set == 1),]
soccer.test <- soccer[which(soccer$set == 2),]
soccer.train$set <- NULL
soccer.test$set <- NULL
soccer.train.x <- soccer.train[,2:length(soccer.train)]
soccer.test.x <- soccer.test[,2:length(soccer.test)]

ridge.info <- c()
ridge.cv.info <- c()

lass.info <- c()
lass.cv.nfo <- c()

#Fit ridge model
ridge_model <- glmnet(as.matrix(soccer.train.x),soccer.train$overall_rating,alpha = 0)

#Calculate MSE and MSPE for ridge model
train.pred <- as.matrix(cbind(const=1,soccer.train.x)) %*% coef(ridge_model)
test.pred <- as.matrix(cbind(const=1,soccer.test.x)) %*% coef(ridge_model)
ridge_model.mse <- mse(train.pred,soccer.train$overall_rating)
ridge_model.mspe <- mse(test.pred,soccer.test$overall_rating)
ridge.info <- c(ridge_model.mse,ridge_model.mspe)

#Fit Lasso model
lasso_model <- glmnet(as.matrix(soccer.train.x),soccer.train$overall_rating,alpha = 1)
train.pred <- as.matrix(cbind(const=1,soccer.train.x)) %*% coef(lasso_model)
test.pred <- as.matrix(cbind(const=1,soccer.test.x)) %*% coef(lasso_model)
lasso_model.mse <- mse(train.pred,soccer.train$overall_rating)
lasso_model.mspe <- mse(test.pred,soccer.test$overall_rating)
lasso.info <- c(lasso_model.mse,lasso_model.mspe)


#Fit model with cross validation
ridge_model <- cv.glmnet(as.matrix(soccer.train.x),soccer.train$overall_rating,alpha = 0,nfolds = 5)
lasso_model <- cv.glmnet(as.matrix(soccer.train.x),soccer.train$overall_rating,alpha = 1,nfolds = 5)

best_lambda.ridge <- ridge_model$lambda.1se
best_lambda.lasso <- lasso_model$lambda.1se
par(mfrow=c(1,2))
plot(ridge_model,main = "Ridge Regression")
plot(lasso_model,main = "Lasso Regression")
ridge_coeff <- ridge_model$glmnet.fit$beta[,ridge_model$glmnet.fit$lambda == best_lambda.ridge]
lasso_coeff <- lasso_model$glmnet.fit$beta[,lasso_model$glmnet.fit$lambda == best_lambda.lasso]

train.pred <- predict(ridge_model,as.matrix(soccer.train.x),s = "lambda.1se")
test.pred <- predict(ridge_model,as.matrix(soccer.test.x),s = "lambda.1se")
ridge_model.cv.mse <- mse(train.pred,soccer.train$overall_rating)
ridge_model.cv.mspe <- mse(test.pred,soccer.test$overall_rating)
ridge.cv.info <- c(ridge_model.cv.mse,ridge_model.cv.mspe)

train.pred <- predict(lasso_model,as.matrix(soccer.train.x),s = "lambda.1se")
test.pred <- predict(lasso_model,as.matrix(soccer.test.x),s = "lambda.1se")
lasso.cv.mse <- mse(train.pred,soccer.train$overall_rating)
lasso.cv.mspe <- mse(test.pred,soccer.test$overall_rating)
lasso.cv.info <- c(lasso.cv.mse,lasso.cv.mspe)

err <- as.table(rbind(ridge.info,lasso.info,ridge.cv.info,lasso.cv.info))
colnames(err) <- c("MSE","MSPE")
rownames(err) <- c("Ridge","LASSO","RidgeCV","LASSOCV")


err
```

```{r,echo=FALSE}
#Compare Coefficients:
coeff <- data.table(Lasso = lasso_coeff,Ridge = ridge_coeff)
coeff[,Features :=names(ridge_coeff)]
to_plot <- melt(data = coeff,id.vars = 'Features',variable.name = 'Model',value.name = 'Coefficients')
 ggplot(to_plot,aes(x=Features,y=Coefficients,fill=Model)) + coord_flip() + geom_bar(stat = 'identity') + facet_wrap(~ Model) + guides(fill = FALSE)
```


## Non-Linear Model
### GAM

### Tensor Producting Smoothing

### Regression Tree

```{r,echo= FALSE}
calErr <- function(tree,data,y) {
  y.hat <- predict(tree, newdata = data)
  return(mse(y,y.hat))
}
```

```{r,echo=FALSE}

rt <- rpart(data = soccer.train, soccer.train$overall_rating ~ . , method = 'anova')
pruneTree.1se <- function(tree) {
  tree$cptable[,c(2:5,1)]
  tree.cpt <- tree$cptable
  minrow <- which.min(tree.cpt[,4])
  se.row <- min(which(tree.cpt[,4] < tree.cpt[minrow,4] + tree.cpt[minrow,5]))
  cplow.1se <- tree.cpt[se.row,1]
  cpup.1se <- ifelse(se.row==1, yes=1, no = tree.cpt[se.row-1,1])
  cp.1se <<- sqrt(cplow.1se*cpup.1se)
  print(cp.1se)
  prune.1se <- prune(tree,cp = cp.1se )
  return(prune.1se)
}

pruneTree.min <- function(tree) {
  tree$cptable[,c(2:5,1)]
  tree.cpt <- tree$cptable
  minrow <- which.min(tree.cpt[,4])
  cplow.min <- tree.cpt[minrow,1]
  cpup.min <- ifelse(minrow==1, yes = 1, no = tree.cpt[minrow-1,1])
  cp.min <<- sqrt(cplow.min * cpup.min)
  prune.mincp <- prune(tree,cp = cp.min)
  return(prune.mincp)
}

prune.1se <- pruneTree.1se(rt)
prune.mincp <- pruneTree.min(rt)


prp(rt, type=1, extra=1, main="Original full tree")

#Calculating Unpruned Regression tree MSE and MSPE
rt.mse <- calErr(rt,soccer.train.x,soccer.train$overall_rating)
rt.mspe <- calErr(rt,soccer.test.x,soccer.test$overall_rating)
rt.info <- c(rt.mse,rt.mspe)
rt.info

#Calculating pruned Regression tree MSE and MSPE
prune.1se.mse <- calErr(prune.1se,soccer.train.x,soccer.train$overall_rating)
prune.1se.mspe <- calErr(prune.1se,soccer.test.x,soccer.test$overall_rating)
prune.1se.info <- c(prune.1se.mse,prune.1se.mspe)

prune.1se.info

prune.mincp.mse <- calErr(prune.mincp,soccer.train.x,soccer.train$overall_rating)
prune.mincp.mspe <- calErr(prune.mincp,soccer.test.x,soccer.test$overall_rating)
prune.mincp.info <- c(prune.mincp.mse,prune.mincp.mspe)

prune.mincp.info


err1 <- as.table(rbind(rt.info,prune.1se.info,prune.mincp.info))
colnames(err1) <- c("MSE","MSPE")
rownames(err1) <- c("RT","1se","min")
```

```{r,echo = FALSE}
par(mfrow=c(1,2))
prp(prune.1se, type=1, extra=1, main="Pruned+1se Tree")
prp(prune.mincp, type=1, extra=1, main="Pruned+mincp Tree")
```


```{r,echo = FALSE}

par(mfrow=c(1,2))
plotcp(prune.1se)
plotcp(prune.mincp)
```


```{r,echo=FALSE}
#Build Regression tree as large as possible
rt.cp0 <- rpart(data = soccer.train, soccer.train$overall_rating ~ ., method = "anova", cp = 0)
par(mfrow=c(1,2))
prune.1se <- pruneTree.1se(rt.cp0)
prune.mincp <- pruneTree.min(rt.cp0)
head(prune.1se$variable.importance)
head(prune.mincp$variable.importance)


rt.cp0.mse <- calErr(rt.cp0,soccer.train.x,soccer.train$overall_rating)
rt.cp0.mspe <- calErr(rt.cp0,soccer.test.x,soccer.test$overall_rating)
rt.cp0.info <- c(rt.cp0.mse,rt.cp0.mspe)



#Calculating pruned Regression tree MSE and MSPE
prune.1se.mse <- calErr(prune.1se,soccer.train.x,soccer.train$overall_rating)
prune.1se.mspe <- calErr(prune.1se,soccer.test.x,soccer.test$overall_rating)
prune.1se.info <- c(prune.1se.mse,prune.1se.mspe)


prune.mincp.mse <- calErr(prune.mincp,soccer.train.x,soccer.train$overall_rating)
prune.mincp.mspe <- calErr(prune.mincp,soccer.test.x,soccer.test$overall_rating)
prune.mincp.info <- c(prune.mincp.mse,prune.mincp.mspe)


err2 <- as.table(rbind(rt.cp0.info,prune.1se.info,prune.mincp.info))
colnames(err2) <- c("MSE","MSPE")
rownames(err2) <- c("RT","1se","min")

```

```{r}
err1
err2
```


### Random Forest

# Statistical Conclusions

# Conclusion in the context of the problem

# Future Work

# Contribution

# Appendix
## Variables
* player_name: The name of the player
* finishing: The accuracy of shots using foot, inside the penalty area
* dribbling: The ability to kepp possesion of the ball.
* ball_control: The ability to keep your ball under your feet with velocity.
* reactions: How quickly a player respinds a situation.
* stamina: Determine the rate at which a player will tire during a game.
* interceptions: The ability to intercepts a pass where the ball is going and stop it from going there.
* marking: The ability to track and defend an opposing player.
* overall_rating: The rating of the player based on all attributes.
* heading_accuracy: The accuracy of he player either a pass or a shot by using head.
* curve: The ability to shoot the ball in a curved shape.
* acceleration: Increase in the rate of speed of a player.
* balance: The ability to maintain balance after a physical challenge.
* strength: The ability to win a physical challenge.
* positioning: The ability to read the game offensively, get into good positions, make effective runs, and avoid getting caught offside.
* standing_tackle:The ability of the player to time standing tackles so that they win the ball rather than give away a foul.
* potential: A peak in overall rating that a player could reach.
* short_passing: The ability to perform a pass in short distance.
* free_kick_accuracy: The accuracy of a direct free kick on goal.(Free kick: an unimpeded kick of the stationary ball awared to one side as a penalty for a foul by the other side)
* sprint_speed: The maximum speed over a short distance of a player.
* shot_power: How hard can the player hits the ball when taking a shot at goal.
* long_shots: The accuracy of shots from outside of the penalty area.
* vision: The player's awareness of the position of his team mates & opponents around him.
* sliding_tackle: The ability of the player to time sliding tackles so that they win the ball rather than give away a foul.
* crossing: The accuracy of the player crosses the ball.
* volleys: The accuracy of a player strike or hit the ball at goal before it touches the ground.
* long_passing: The ability to perform a long pass in the air and on the ground to his teammate.
* agility: The ability of a player to move or turn in game.
* jumping: The vertical distance of a player can jump from the ground.
* aggression: The frequency & aggression of jostling, tackling & slide tackling.
* penalties: The ability to take penalties.
* gk_reflexes: The ability to react a ball in movement at goal by the goal keeper.


## R-Code
```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```
