---
title: "Final_Project_r24tan"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/Raymond/Desktop/Raymond Tan/HW/4B/STAT444/soccer-rating-prediction/data")
soccer.raw <- read.table("rating_potential.csv",sep = " ",na.strings = "NA")

```


```{r}
colnames(soccer.raw)
```

* player_name: The name of the player
* finishing: The accuracy of shots using foot, inside the penalty area
* dribbling: The ability to kepp possesion of the ball.
* ball_control: The ability to keep your ball under your feet with velocity.
* reactions: How quickly a player respinds a situation.
* stamina: Determine the rate at which a player will tire during a game.
* interceptions: The ability to intercepts a pass where the ball is going and stop it from going there.
* marking: The ability to track and defend an opposing player.
* overall_rating: The rating of the player based on all attributes.
* heading_accuracy: The accuracy of he player either a pass or a shot by using head.
* curve: The ability to shoot the ball in a curved shape.
* acceleration: Increase in the rate of speed of a player.
* balance: The ability to maintain balance after a physical challenge.
* strength: The ability to win a physical challenge.
* positioning: The ability to read the game offensively, get into good positions, make effective runs, and avoid getting caught offside.
* standing_tackle:The ability of the player to time standing tackles so that they win the ball rather than give away a foul.
* potential: A peak in overall rating that a player could reach.
* short_passing: The ability to perform a pass in short distance.
* free_kick_accuracy: The accuracy of a direct free kick on goal.(Free kick: an unimpeded kick of the stationary ball awared to one side as a penalty for a foul by the other side)
* sprint_speed: The maximum speed over a short distance of a player.
* shot_power: How hard can the player hits the ball when taking a shot at goal.
* long_shots: The accuracy of shots from outside of the penalty area.
* vision: The player's awareness of the position of his team mates & opponents around him.
* sliding_tackle: The ability of the player to time sliding tackles so that they win the ball rather than give away a foul.
* crossing: The accuracy of the player crosses the ball.
* volleys: The accuracy of a player strike or hit the ball at goal before it touches the ground.
* long_passing: The ability to perform a long pass in the air and on the ground to his teammate.
* agility: The ability of a player to move or turn in game.
* jumping: The vertical distance of a player can jump from the ground.
* aggression: The frequency & aggression of jostling, tackling & slide tackling.
* penalties: The ability to take penalties.
* gk_reflexes: The ability to react a ball in movement at goal by the goal keeper.






